
# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
```{r}
library(data.table)
library(plyr)
library(lubridate)
data <- fread("activity.csv", colClasses=c("numeric","factor","integer"))
data$date <- ymd(data$date)
```
## What is mean total number of steps taken per day?
```{r}
total_steps <- data[, list( sum = sum(steps)), by = date]

# -------------------------------------------------
# ALTERNATIVE

# total_steps <- tapply(data$steps, data$date, sum)
# -------------------------------------------------

hist(total_steps$sum, cex.lab = 1.4, xlab = "Total steps per day", main = "Histogram of the total number of steps taken each day")

# mean and median total number of steps taken per day
mean(total_steps$sum, na.rm = T)
median(total_steps$sum, na.rm = T)
```
## What is the average daily activity pattern?
```{r}
head(data[, avg := mean(steps, na.rm = T), by = interval], 5)

# ----------------------------------------------------------
# SIMILAR CODE LIKE PLYR ddply

# data <- data[, list(avg = mean(steps, na.rm = T)), by = interval]
# -----------------------------------------------------------

plot(avg ~ interval, data, type = "l", cex.lab = 1.4, xlab = "Intervals", ylab = "Average number of steps")

# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

data[order(-avg, interval), avg, interval][1]

# -------------------------------------------------
#  ALTERNATIVE IN PLYR

#  arrange(data[ ,avg, interval], desc(avg))[1]
# ------------------------------------------------- 

```
## Imputing missing values
```{r}
#  missing values
sum(is.na(data))

# Imputing NA's with average on 5-min interval and create new dataset
head(data[is.na(steps), steps := avg], 5)
daily_steps <- data[, list(sum = sum(steps)), by = date]

# ------------------------------------------------
# daily_steps <- tapply(data$steps, data$date, sum)
# ------------------------------------------------

# plot the histogram
hist(daily_steps$sum, cex.lab = 1.4, xlab = "Total steps per day", main = "Histogram of the total number of steps per day")

# mean and median total number of steps taken per day don't change significantly
mean(daily_steps$sum)
median(daily_steps$sum)

# difference between mean and median from the first and this assignment
abs(mean(daily_steps$sum) - mean(total_steps$sum, na.rm = T))
abs(median(daily_steps$sum) - median(total_steps$sum, na.rm = T))
```
Mean values didn't change as imputation and median change slightly.
Imputation change the apperance of histogram. The intervals between
10000-15000 are more frequent.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
# change local time
Sys.setlocale("LC_TIME", "C")
# change date to weekdays
head(data[, date := weekdays(date, abbreviate = T)], 5)
# create string column week with two variable "weekend" and "weekday" by ifelse condition
head(data[, date := ifelse(date %in% c("Sat", "Sun"), "WeekEnd", "WeekDay")], 5)
# summarize steps mean by interval and date - this is plyr like
week <- data[, list(steps = mean(steps)), by = c("interval", "date")]

# ----------------------------------------------------------------------
#  ALTERNATIVE
#  week <- ddply(data, .(interval, date), summarize, steps = mean(steps))
# ----------------------------------------------------------------------

# create plot
xyplot(steps ~ interval | date, week, layout = c(1, 2), xlab =
       list("Intervals", cex = 1.3), ylab = list("Steps", cex = 1.3), type = "l")



```
